---
title: "Analysis of finger/symbol experiment"
author: "Alexander D Forrence"
date: "September 28, 2015"
output: html_document
---


```{r, message = FALSE, warning = FALSE}
setwd("C:/Users/aforr_000/Desktop/lab-materials/force-analysis")


library(arm)
library(boot)
library(dplyr)
library(ggplot2)
library(nlme)
library(psyphy)
library(zoo)

blks <- list()
blknames <- list.files('csvfiles')
i <- 1
for (file in blknames) {
    blks[[i]] <- read.csv(paste0('csvfiles/',file), header = TRUE)
    i <- i + 1
}

#dplyr version
all_blks <- blks %>%
    bind_rows() %>%
    na.omit() %>%
    mutate(success = ifelse(targ_finger == press_finger, 1, 0)) %>%
    mutate(RT = t_press - t_stim) %>%
    filter(RT > 0) %>%
    add_rownames(var = 'ind') %>%
    mutate(ind = as.numeric(ind)) %>%
    arrange(RT) %>%
    mutate(cond = ifelse(t_stim == 0.75, 'prac', 'var'))

```

## Looking at the raw data

```{r, fig.height = 5,echo=FALSE}
ggplot(all_blks, aes(x = ind, y = RT, color = factor(success))) + 
    geom_point(alpha = .6) + 
    facet_wrap(~day, scales = "free_x")
```

```{r, fig.height = 5,echo=FALSE}
ggplot(all_blks, 
       aes(x = RT, y = factor(targ_finger), 
           color = factor(press_finger))) + 
    geom_jitter(alpha = .6) + 
    facet_wrap(~day)

```

The `glm` function uses iteratively reweighted least squares (IRLS) to find a maximum likelihood solution.

```{r,warning = FALSE,message = FALSE}
df <- filter(all_blks, cond == "var" & day > 1)
df_roll <- data.frame(roll_succ = rollmean(df$success, 70),
                      roll_rt = rollmean(df$RT, 70))
m0 <- glm(success ~ RT,
          data = df,
          family = binomial(psyphy::mafc.probit(4)))
m2 <- update(m0, . ~ 0 + RT * factor(day))
```

For some reason, profiling the likelihood doesn't work properly. Maybe `profile.glm` from MASS doesn't deal with "custom" link functions properly?

```{r}
summary(m2)
```

RT on the x axis, P(success) on the y, fit cumulative gaussian

TODO: fit cumulative gaussian w/ boundaries (via mle2 in bbmle?)

## Examining the effect of days (could we do it block-by-block?)

Below are the "raw" data, along with the success plotted by reaction time.

```{r,echo=FALSE}

df$resp2 <- predict(m2, type = "response")

df_roll <- list()
days <- 2:5
for (i in 1:4) {
    tempdf <- df %>% 
        filter(day == days[i]) %>%
        arrange(RT)
    
    df_roll[[i]] <- data.frame(
        success = rollmean(tempdf$success, 70),
        RT = rollmean(tempdf$RT, 70),
        day = days[i]
    )
}

df_roll <- bind_rows(df_roll)

ggplot(df, 
       aes(RT, press_finger, 
           colour = factor(targ_finger))) + 
    geom_jitter(size = 2.3, width = .1,
                height = .2, alpha = .6) +
    labs(x = "Reaction Time (s)", 
         y = "Target Finger",
         colour = "Digit\ndigit")

ggpsych <- ggplot(data = df, aes(x = RT, y = success)) + 
    geom_line(aes(y = resp2, 
                  colour = "factor(days) GLM"),
              size = 1.5, alpha = 0.6) + 
    geom_line(data = df_roll, 
              aes(RT, success, 
                  colour = "Rolling mean"), 
              size = 1.5) +
    geom_point(aes(colour = "Raw Values"), size = 2, shape = 124) +
    facet_wrap(~day)

ggpsych
```

## Autocorrelation in the response?

```{r,echo=FALSE}
df2 <- df %>% arrange(ind) #rearrange by index
acf(df2$press_finger, ci.type = "ma")
    
```

But there doesn't seem to be much in the response residuals of the model (maybe due to the lack of data?)

```{r,echo=FALSE}

acf(resid(m2, type = "response"))

```

## Bootstrapping the midpoint

```{r, message = FALSE, warning = FALSE}

glmfit <- function(data,indices) {
    fit <- glm(success ~ 0 + RT * factor(day), 
               data = data[indices,], 
               family = binomial(psyphy::mafc.probit(4)))
    temp <- coef(fit)
    # -b0/b1 for each day
    b <- c(0, 0, 0, 0)
    b[1] <- -temp[2] / temp[1]
    b[2] <- -temp[3] / (temp[1] + temp[6])
    b[3] <- -temp[4] / (temp[1] + temp[7])
    b[4] <- -temp[5] / (temp[1] + temp[8])
    return(b)
}

set.seed(1)

bb <- boot(data = df, statistic = glmfit, R = 1000)

boots <- t(sapply(X = 1:4,
                  FUN = function(x) boot.ci(bb, index = x, type = "perc")$percent))
boots <- data.frame(day = 2:5, orig = bb$t0, 
                    lowci = boots[,4], hici = boots[,5])
boots

```

```{r,echo=FALSE}
zz <- ggplot(boots, 
       aes(x = day, 
           y = orig, 
           colour = factor(day))) + 
    geom_errorbar(aes(ymin = lowci,ymax = hici),
           size = 1, width = .1, alpha = .5) +
    geom_point(size = 4) + 
    theme_bw() + 
    ylab("Mean (s)")
zz
ggpsych <- ggpsych + geom_vline(data = boots, 
                     aes(xintercept = orig), 
                     size = 1, alpha = .4)
ggpsych
```

Using `stats::nls()`, which will give us something that looks more similar to the maximum likelihood:

```{r}

df <- df %>% arrange(day, RT)

coefs <- list(mean = .4, sd = .1)
mods <- nlsList(success ~ 1/4 + (1 - 1/4) * pnorm(RT, mean, sd)| day, 
         data = df, 
         start = coefs)


df <- df %>% 
    arrange(day, RT) %>% 
    mutate(prednls = c(predict(mods, df)))

ggplot(data = df, aes(x = RT, y = success)) + 
    geom_line(aes(y = resp2, 
                  colour = "factor(days) GLM"),
              size = 1.5, alpha = 0.6) + 
    geom_line(data = df_roll, 
              aes(RT, success, 
                  colour = "Rolling mean"), 
              size = 1.5) +
    geom_point(aes(colour = "Raw Values"),
               size = 2, shape = 124) + 
    geom_line(data = df, 
              aes(x = RT, y = prednls, color = "NLS"),
              size = 1.5, linetype = "dashed")+
    facet_wrap(~day)
   
```

Which gives pretty similar answers to the `glm`.

```{r, echo = FALSE}
summary(mods)
```

And profile confidence intervals on the mean:

```{r, echo = FALSE}

nlsconf <- data.frame(
    t(sapply(X = mods, FUN = confint)[c(1,3,2,4),])
)

temp <- t(sapply(X = mods, FUN = coef))

colnames(nlsconf) <- c("mean05", "mean95",
                       "sd05", "sd95")
nlsconf$meancoef <- temp[,1]
nlsconf$sdcoef <- temp[,2]
nlsconf$day <- 2:5
nlsconf <- nlsconf[,c(7,5,6,1,2,3,4)]

ggplot(boots, 
       aes(x = day, 
           y = orig, 
           colour = "Bootstrapped GLM Est.")) + 
    geom_errorbar(aes(ymin = lowci,ymax = hici),
           size = 1, width = .1, alpha = .5) +
    geom_point(size = 4) + 
    theme_bw() + 
    ylab("Mean (s)") + 
    geom_point(data = nlsconf, 
               aes(x = day + .15, y = meancoef,
                   colour = "Profiled NLS Est."),
               size = 4) + 
    geom_errorbar(data = nlsconf, 
                  aes(x = day + .15, y = NULL,
                      ymax = mean95, ymin = mean05,
                   colour = "Profiled NLS Est."), 
                  size = 1, width = .1)
```

I also didn't work out the equivalent of `sd` in the bootstrapped version, but the profile nls can be found below.

```{r, echo = FALSE}

ggplot(nlsconf, aes(x = day, y = sdcoef, color = factor(day))) + 
    geom_point(size = 4) + 
    geom_errorbar(aes(ymin = sd05, ymax = sd95), size = 1, width = .1) + 
    theme_bw()

```

It looks like Michael leveled out (more or less) by day 3 - 4.